---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r}
#| label: pkg_load
#| echo: false
#| message: false
#| warning: false
require(tidywigits)
```

# 🧬✨ tidywigits

- Docs: <https://umccr.github.io/tidywigits/>

{tidywigits} is an R package that parses and tidies outputs from the [WiGiTS]
suite of genome and transcriptome analysis tools for cancer research and diagnostics,
created by the Hartwig Medical Foundation.

[WiGiTS]: <https://github.com/hartwigmedical/hmftools> "WiGiTS suite"

In short, it traverses through a directory containing results from one or more
runs of WiGiTS tools, parses any files it recognises, tidies
them up (which includes data reshaping, normalisation, column name cleanup etc.),
and writes them to the output format of choice e.g. Apache Parquet, PostgreSQL,
Delta table.

## 🍕 Installation

### R

```r
remotes::install_github("umccr/tidywigits")
```

### Conda

#### Linux & MacOS (non-ARM)

```bash
conda create \
  -n tidywigits_env \
  -c umccr -c conda-forge \
  r-tidywigits==X.X.X

conda activate tidywigits_env
```

#### MacOS ARM

```bash
CONDA_SUBDIR=osx-64 \
  conda create \
  -n tidywigits_env \
  -c umccr -c conda-forge \
  r-tidywigits==X.X.X

conda activate tidywigits_env
```

### Docker

```bash
docker pull --platform linux/amd64 ghcr.io/umccr/tidywigits:X.X.X
```

## 🌀 CLI

A `tidywigits.R` command line interface is available for convenience.

- If you're using the conda package, the `tidywigits.R` command will already be
  available inside the activated conda environment.
- If you're _not_ using the conda package, you need to export the
  `tidywigits/inst/cli/` directory to your `PATH` in order to use `tidywigits.R`.

```{bash}
#| label: cli1
#| echo: true
#| eval: false
tw_cli=$(Rscript -e 'x = system.file("cli", package = "tidywigits"); cat(x, "\n")' | xargs)
export PATH="${tw_cli}:${PATH}"
```

```{bash}
#| label: cli2
#| echo: false
#| prompt: true
tw_cli=$(Rscript -e 'x = system.file("cli", package = "tidywigits"); cat(x, "\n")' | xargs)
export PATH="${tw_cli}:${PATH}"

echo "tidywigits.R --version" & tidywigits.R --version
echo ""
echo "#-----------------------------------#"
echo "tidywigits.R --help" & tidywigits.R --help
echo ""
echo "#-----------------------------------#"
echo "#------- Tidy ----------------------#"
echo "tidywigits.R tidy --help" & tidywigits.R tidy --help
echo ""
echo "#-----------------------------------#"
echo "#------- List ----------------------#"
echo "tidywigits.R list --help" & tidywigits.R list --help
echo ""
```