---
title: "NEWS"
vignette: >
  %\VignetteIndexEntry{News}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

```{r}
#| label: pkgs
#| message: false
#| warning: false
#| echo: false
{
  use("htmltools", "a")
  use("glue", "glue")
}
```

```{r echo=FALSE}
repo <- "https://github.com/umccr/tidywigits"
vdiff <- function(new, old) {
  stopifnot(utils::compareVersion(new, old) == 1)
  htmltools::a(
    href = glue::glue("{repo}/compare/v{old}...v{new}"),
    glue::glue("v{new} - v{old} diff")
  )
}

pr_or_issue <- function(n, type) {
  x <- c(iss = "issues", pr = "pull")
  stopifnot(is.numeric(n), n > 0, type %in% names(x))
  type2 <- x[type]
  htmltools::a(
    href = glue::glue("{repo}/{type2}/{n}"),
    glue::glue("{type}{n}")
  )
}

iss <- function(n) {
  pr_or_issue(n, "iss")
}

pr <- function(n) {
  pr_or_issue(n, "pr")
}
```

## DEV

## v0.0.3 (2025-08-04)

- `r vdiff("0.0.3", "0.0.2")`

Fixed bug where `normalizePath` was getting called with NULL output directory in the db format case (`r pr(133)`).

## v0.0.2 (2025-07-14)

- `r vdiff("0.0.2", "0.0.1")`

Mostly Shiny, Conda, Docker, pkgdown and GitHub Actions support.

- Add GitHub Actions for deployment of the following (`r pr(130)`, `r iss(9)`):
  - Conda: add recipe and rattler-builder (`r pr(124)`, `r iss(4)`)
  - Docker: add Dockerfile (`r pr(130)`)
  - Shiny: add summary app (`r pr(122)`)
  - CLI: Add listing and tidy include/exclude support (`r pr(116)`)
  - Add pkgdown support (`r pr(125)`, `r iss(37)`)
  - Add DVC support with some purple test data (`r pr(126)`, `r iss(59)`)
- Optimise file listing (`r pr(131)`, `r iss(127)`)
  - Remove `File` class

## v0.0.1 (2025-06-19)

Initial release of tidywigits.

- R pkg skeleton (`r pr(6)`)
- Add `Config`, `File` and `Tool` classes (`r pr(12)`)
- Add `Workflow` class (`r pr(99)`, `r iss(97)`)
- Add `Oncoanalyser` class (`r pr(52)`, `r pr(101)`, `r iss(39)`)
- Add `bump-my-version` (`r pr(10)`, `r iss(3)`)
- Add `pre-commit` hooks (`r pr(8)`, `r iss(2)`)
- Add `Makefile` and `air.toml` (`r pr(7)`)
- Support for main outputs from the following WiGiTS tools:
  - Alignments (`r pr(28)`, `r pr(79)`, `r iss(17)`, `r iss(77)`)
  - Amber (`r pr(14)`, `r iss(13)`)
  - Bamtools (`r pr(44)`, `r pr(66)`, `r iss(62)`, `r iss(16)`)
  - Chord (`r pr(29)`, `r pr(38)`, `r iss(18)`, `r iss(38)`)
  - Cobalt (`r pr(33)`, `r iss(15)`)
  - Cuppa (`r pr(30)`, `r pr(68)`, `r iss(19)`, `r iss(67)`)
  - Esvee (`r pr(87)`, `r iss(61)`)
  - Flagstats (`r pr(45)`, `r pr(46)`, `r iss(20)`)
  - Isofox (`r pr(80)`, `r iss(76)`)
  - Lilac (`r pr(31)`, `r iss(21)`)
  - Linx (`r pr(50)`, `r pr(89)`, `r iss(22)`, `r iss(88)`)
  - Purple (`r pr(36)`, `r pr(51)`, `r iss(23)`, `r iss(49)`)
  - Sage (`r pr(48)`, `r pr(72)`, `r pr(73)`, `r iss(24)`, `r iss(71)`, `r iss(57)`)
  - Sigs (`r pr(47)`, `r iss(25)`)
  - Virusbreakend (`r pr(42)`, `r iss(26)`)
  - Virusinterpreter (`r pr(40)`, `r pr(65)`, `r iss(27)`, `r iss(41)`)
- Schema: 
  - versioning (`r pr(64)`, `r iss(43)`)
  - guesser (`r pr(69)`, `r iss(63)`)
- Vignettes:
  - Setup (`r pr(92)`, `r iss(60)`)
- Tool:
  - write functionality (`r pr(98)`, `r iss(95)`)
  - optionally output raw parsed tibbles (`r pr(94)`, `r iss(91)`)
- UML diagram (`r pr(107)`, `r iss(106)`)
- DB schema (`r pr(111)`, `r iss(110)`)
- CLI support (`r pr(114)`)
